### 2012-10-19 thesis/MA-Thesis.tex
* increased font size, from 10 to 12pt * increased linespacing, now: 1,5 * Chapter "Design" is now splittet into "Visual formalisms" and   "Automated constraint generation" * Section "6.1 Operator constraints" now filled (with lots of pictures) * smaller changes due to new font size and linespacing

\documentclass[12pt,a4paper,headsepline,footsepline,DIV13,BCOR12mm]{scrbook}
\usepackage[english,ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{scrpage2}
\pagestyle{scrheadings}

\usepackage{graphicx}

\usepackage{setspace}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{amsmath}

% LTL Symbols
\usepackage{latexsym}
\usepackage{MnSymbol}

\usepackage{wrapfig}

\usepackage{listings}
\usepackage{color}
\definecolor{gray}{rgb}{0.4,0.4,0.4}
\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}

\begin{document}

%PREFACE
\selectlanguage{ngerman}
\input{titlepage}

\onehalfspacing
\newpage
\thispagestyle{empty}

\centerline{\bfseries ERKLÄRUNG}

\vspace{5cm}
Hiermit versichere ich, dass ich diese Masterarbeit selbständig verfasst habe.
Ich habe dazu keine anderen als die angegebenen Quellen und Hilfsmittel
verwendet.

\vspace{1cm}
\begin{flushleft}
%select german for formatting the date
\selectlanguage{ngerman}
Augsburg, den \today \hfill Ludwig Nägele
\end{flushleft}

\newpage
\thispagestyle{empty}
\mbox{}

\newpage

%select english as language!
\selectlanguage{english}
%\pagenumbering{roman}
\pagenumbering{arabic}
\tableofcontents
\newpage
\listoffigures
\newpage
\listoftables
%\pagenumbering{arabic}

\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.jpg}

\input{tex/1_abstract}
\input{tex/2_introduction}
\input{tex/3_background}
\input{tex/4_goals}
\input{tex/5_design}
\input{tex/6_implementation}
\input{tex/7_prototypeandevaluation}
\input{tex/8_conclusionandfuturework}

\chapter*{Acknowledgment}

%This work evolved from a master thesis project provided by the Elite Graduate Program Software Engineering. 
%For the received support I would like to thank the people of this program and the three patron universities: University of Augsburg, Technical University Munich and Ludwig-Maximilians-University Munich.

\bibliographystyle{unsrt}
\bibliography{Bib}

\end{document}