### 2012-10-21 thesis/MA-Thesis.tex
The .tex file structure and names have been changed. Now they match the numbers and titles in the table of contents.

\documentclass[12pt,a4paper,headsepline,footsepline,DIV13,BCOR12mm]{scrbook}
\usepackage[english,ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{scrpage2}
\pagestyle{scrheadings}

\usepackage{graphicx}

\usepackage{setspace}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{amsmath}

% LTL Symbols
\usepackage{latexsym}
\usepackage{MnSymbol}

\usepackage{wrapfig}

\usepackage{listings}
\usepackage{color}
\definecolor{gray}{rgb}{0.4,0.4,0.4}
\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}

%for bookmark bar
\usepackage[colorlinks,pdfpagelabels,pdfstartview=FitH,bookmarksopen=false,bookmarksnumbered=false,linkcolor=black,plainpages=false,hypertexnames=false,citecolor=black]{hyperref}

\usepackage{algorithm}
\usepackage{algorithmic}

\begin{document}

%PREFACE
\selectlanguage{ngerman}
\input{titlepage}

\onehalfspacing
\newpage
\thispagestyle{empty}

\centerline{\bfseries ERKLÄRUNG}

\vspace{5cm}
Hiermit versichere ich, dass ich diese Masterarbeit selbständig verfasst habe.
Ich habe dazu keine anderen als die angegebenen Quellen und Hilfsmittel
verwendet.

\vspace{1cm}
\begin{flushleft}
%select german for formatting the date
\selectlanguage{ngerman}
Augsburg, den \today \hfill Ludwig Nägele
\end{flushleft}

\newpage
\thispagestyle{empty}
\mbox{}

\newpage

%select english as language!
\selectlanguage{english}
%\pagenumbering{roman}
\pagenumbering{arabic}

\pdfbookmark[1]{Inhaltsverzeichnis}{toc}
\tableofcontents
\newpage
\listoffigures
\newpage
%\listoftables
\lstlistoflistings
%\pagenumbering{arabic}

\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.jpg}

\input{tex/0_abstract}
\input{tex/1_introduction}
\input{tex/2_background}
\input{tex/3_goals}
\input{tex/4_visualformalisms}
\input{tex/5_automatedconstraintgeneration}
\input{tex/6_theltlcreatorprototype}
\input{tex/7_testscenarioandevaluation}
\input{tex/8_conclusionandfuturework}

\chapter*{Acknowledgment}

%This work evolved from a master thesis project provided by the Elite Graduate Program Software Engineering. 
%For the received support I would like to thank the people of this program and the three patron universities: University of Augsburg, Technical University Munich and Ludwig-Maximilians-University Munich.

\bibliographystyle{unsrt}
\bibliography{Bib}

\end{document}