### 2012-10-21 thesis/tex/0_abstract.tex

\chapter*{Abstract}

Nowadays robotic applications are no longer confined to only industrial use but increasingly assigned to jobs within human workspaces. Even the programming of robots is shifting from experienced software developers to the consumers themselves who need to customize the robot's behavior in order to match individual requirements.
However both the human robot collaboration as well as the behavior definition by non experts may be a risk to humans within the robots' environment.
Thus special thinking is needed about how to cope with safety requirements for robot behavior specification by consumers.

In this paper we present a visual language for the defining of safety constraints for state machine definition of robot behaviour. Our approach addresses mainly non-safety-experts and our abstraction from a mathematical temporal logic expression to a more intuitive visual representation is intended to enable a wider range of software developers to do model checking.
Furthermore we propose a new concept of development support by providing automated constraint generation functionality. It aims to help developers in defining reasonable constraints and finally to increase the probability of finding bugs.

A graphical editor has been developed which demonstrates both visual constraint creation functionality as well as automated constraint generation.