### 2012-11-13 thesis/tex/0_abstract.tex

\chapter*{Abstract}

Nowadays robotic applications are no longer confined to only industrial use but increasingly assigned to tasks within human workspaces and living environments. Even the programming of robots is shifting from experienced software developers to the consumers themselves who need to customize the robot's behaviour in order to match individual requirements.
However, both human robot collaboration as well as the robot's behaviour definition by non experts may be a risk to humans within the robots' environment.
Thus, special thinking is needed about how to cope with safety requirements for robot behaviour specification by consumers.

In this work a visual language for the defining of safety constraints for state machine definition of robot behaviour is presented. The approach addresses mainly non-safety-experts and introduces abstraction from a mathematical temporal logic expression to a more intuitive visual representation to enable a wider range of software developers to do model checking.
Furthermore, a new concept of development support by providing automated constraint generation functionality is proposed. It aims to help developers in defining reasonable constraints and finally to increase the probability of finding bugs.

A graphical editor has been developed which integrates and demonstrates both visual constraint creation functionality as well as automated constraint generation and makes them usable in an uniform manner.